---
title: "Data Preparation"
author: "Bobby Nelson"
date: "2/3/2022"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the Full Data

The data for this application is the CDC's published data on United States
Chronic Disease Indicators (CDI) found [here](https://catalog.data.gov/dataset/u-s-chronic-disease-indicators-cdi).
This dataset provides a detailed breakdown of U.S. chronic disease indicators 
by state. There is a lot of extra detail in the data that should be simplified
before loading into the Shiny application. You can read more about the data
[here](./rr6401.pdf).

```{r}
full.data <- read_csv("U.S._Chronic_Disease_Indicators__CDI_.csv")
```

# Subset the Data

We only want a small subset of the features in the data.

```{r}
subset.data <- full.data %>%
  select(LocationAbbr,
         Topic,
         Question,
         DataValue,
         DataValueType,
         DataValueUnit,
         StratificationCategory1,
         Stratification1,
         YearStart)

subset.data <- subset.data %>%
  mutate(state = factor(LocationAbbr),
         Topic = factor(Topic),
         Question = factor(Question),
         DataValue = as.numeric(DataValue),
         DataValueType = factor(DataValueType),
         DataValueUnit = factor(DataValueUnit),
         StratificationCategory1 = factor(StratificationCategory1),
         Stratification1 = factor(Stratification1),
         YearStart = factor(YearStart))

subset.data <- subset.data %>%
  filter(DataValueType == "Crude Rate",
         grepl("Mortality|mortality", Question),
         StratificationCategory1 %in% c("Overall", "Gender"))

subset.data <- subset.data %>%
  mutate(DataValue = ifelse(DataValueUnit=="cases per 100,000",
                            DataValue * 10,
                            DataValue),
         DataValueUnit = "cases per million")
  
```

# Save the Subset

Now save the subset data for the application to use.

```{r}
write_csv(subset.data, "../app/cdi_data.csv")
```

```{r}
saveRDS(subset.data, "../app/cdi_data.rds")
```

```{r}
subset.data %>%
  filter(Topic == "Alcohol") %>%
  group_by(Topic, state, Stratification1) %>%
  summarize(total = sum(DataValue, na.rm=TRUE))
```


